\racets: Policy{-}Agnostic Programming in Racket

### 2018-07-24 paper/paper.tex
making code smaller, I think..?
draft of conclusion

\usepackage{booktabs} % For formal tables
\usepackage[linguistics]{forest}
\usepackage{tikz}
\usetikzlibrary{cd}
\usepackage{xcolor}
\usepackage[labelsep=quad,indention=10pt]{subfig}
\usepackage{mathpartir} % Inference rules
\usepackage[ruled]{algorithm2e} % For algorithms
\usepackage{setspace}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

@@

Because of Racket's flexible macro system, our implementation of
faceted execution is much smaller than other systems: our core macros
comprise roughly 170 source lines of Racket, with another 120 lines of
library code to perform various facet-related functions.

\paragraph*{Case Study: Battleship in Racets}

\begin{figure*}
  \subfloat[]{%
    \includegraphics[height=1.1in]{a.png}%
    \label{fig:player1}%
  } 
  \quad
  \subfloat[]{%
    \includegraphics[height=1.1in]{b.png}%
    \label{fig:player1player2}%
  }
  \quad
  \subfloat[]{%
    \includegraphics[height=.7in]{c.png}%
    \label{fig:player2strike}%
  }
  \vspace{.3in}
  \quad
  \subfloat[]{%
    \includegraphics[height=1.1in]{d.png}%
    \label{fig:player1afterstrike}%
  }
  \caption{Screenshots from our Battleship case study.}
  \label{fig:case}
\end{figure*}

To gain perspective on how Racets enables policy-agnostic programming,
we scaled our implementation from Section~\ref{sec:overview} to a
web-based game of Battleship written in Racets. Our implementation
uses Racket's \texttt{web-server} framework~\cite{webserver}, which
defines an API for writing HTTP-based server applications.

Figure~\ref{fig:case} shows several screenshots of our Battleship
application. The first screenshot~\ref{fig:player1} shows the board as
viewed by player 1 (note the parameter \texttt{/player1/player1}),
while the second shows the empty board observed when player 2 attempts
to view player 1's board. The screenshot in~\ref{fig:player2strike}
shows the response seen by player 2 upon a successful hit. Finally
screenshot~\ref{fig:player1afterstrike} shows player one's board with
the ship on \bmth{(3,2)} removed.

When running, our game server provides several routes that a user can
access:

\begin{lstlisting}[language=Racket,escapechar=|,name=case]
(define-values (dispatch generate-url)
  (dispatch-rules
    [("player1" (string-arg)) player1board]
    [("player2" (string-arg)) player2board]
    [("player1strike" (string-arg)) p1strike]
    [("player2strike" (string-arg)) p2strike]))
\end{lstlisting}

The route \texttt{/player1/<id>} (or \texttt{/player2/<id>}) renders
player 1's game board when viewed as \texttt{<id>}. We facet game
boards with policies that reveal player 1's board when \texttt{<id>}
is \texttt{player1}, and do the same with player 2:

\begin{lstlisting}[language=Racket,escapechar=|,name=case]
(define p1l (mkpol "player1"))
(define p1board
  (box (fac p1l
            (add-pieces (makeboard) '(1 2 ldots))
            (makeboard))))
\end{lstlisting}

The board is explicitly made into a box: this is because the board's
state will change as player 2 makes moves and eliminates pieces from
their board. The \code{player1board} function implements the logic to
render player 1's board as an HTML table. The argument \code{viewer}
corresponds to the \texttt{<id>} route argument, and is passed to
\code{player1board} by the framework. This argument is then used to
observe the board game

\begin{lstlisting}[language=Racket,escapechar=|,name=case]
(define player1board
  (ext-lambda (request viewer)
         (http-response "<h1>Player 1's Game Board</h1>"
                        (pretty-print 
                           (obs p1l name (deref p1board))))))
\end{lstlisting}

The implementation of \code{player1board} uses a special form
\code{ext-lambda}, discussed at the end of this section, to allow code
from Racets to be executed by the framework (which is not prepared to
execute faceted code).

The function \code{p1strike} allows player 1 to make a guess as to the
position of ships on player 2's board. The function parses the
position into two coordinates and then calls \code{mark-hit} to
perform the hit, updating player 2's board and then observing the
result to answer (to player 1) whether the result was a hit or not:

\begin{lstlisting}[language=Racket,escapechar=|,name=case]
(define p1strike
  (ext-lambda
   (request position)
   (let* ([x (char-to-num (string-ref position 0))]
          [y (char-to-num (string-ref position 2))]
          [ans (mark-hit p2board x y)]
      (set! p2board (car ans))
      (http-response
       (if (cdr (obs p2l "player2" ans)) |label{lineno:explicitobs}|
           "<h1>Congratulations!</h1> <h4>You hit player 2!</h4>"
           "<p>No hit :(</p>")))))))
\end{lstlisting}

Note that we need to use an explicit \code{obs} form on
line~\ref{lineno:explicitobs}. This is because---as \code{p2board} is
a facet---the result will also be a faceted value. When we want to
display the output to player 1, our code needs to explicitly observe
the answer, as \code{http-response} cannot accept a faceted value.

\paragraph*{}

\section{Conclusion}
\label{sec:conclusion}

asdf asdf

>>>>

\usepackage{booktabs} % For formal tables
\usepackage[linguistics]{forest}
\usepackage{tikz}
\usetikzlibrary{cd}
\usepackage{xcolor}
\usepackage{mathpartir} % Inference rules
\usepackage[ruled]{algorithm2e} % For algorithms
\usepackage{setspace}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

@@

Because of Racket's flexible macro system, our implementation of
faceted execution is much smaller than other systems: our core macros
comprise only 170 source lines of Racket, with another 120 lines of
library code to perform various facet-related functions.

We plan to make our prototype of \racets available publicly.

\section{Conclusion and Future Work}
\label{sec:conclusion}

In this paper, we have reviewed the operation of faceted execution, a linguistic paradigm
enabling policy-agnostic programming, and showed how it may be implemented within
the Racket programming system as a library of macros. As Racket macros permit core language
forms (including function application, $\lambda$-abstraction, conditionals, mutation, etc.)
to be rewritten arbitrarily, it is possible to rewrite the meaning of these forms to support
a faceted semantics directly. We call our prototype system \racets: Racket with Facets.

The advantage of this approach is that faceted, policy-agnostic, programs may be written
directly in Racket, making use of the wealth of Racket code already available.
A central challenge of this then, is how to ensure there is a sound and practical inter-operation
between racets and standard Racket (or other languages written as libraries in Racket). Our
approach to this has been to use a tagging scheme that identifies values from \racets
so untagged values may be treated by \racets as originating from a non-\racets language.
For example, a pure Racket function that is not tagged, being applied at a $\blue{\code{ppapp}}$
form in \racets, can be automatically lifted to support FE (so that it can split when applied
on a faceted value).

Our hypothesis is that this tagging scheme is key to permitting inter-operation
between \racets and Racket, and we have implemented a faceted, web-based game of Battleship
to explore this idea. We suspect that a more thorough investigation of likely idioms
for faceted, non-faceted module interaction is needed and that purely functional code
plays a special role as a degenerate case where arbitrary non-faceted code may be lifted
to operate over facted values without potential unsoundness. In the future, we plan to explore
these design choices in a more principled way and also to scale a static analysis of faceted
execution to fully expanded Racket so it may be applied directly to \racets.
